Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2019-02-07 paper/ms.tex
Added a description of the flame model to Carlyn's paper

\section{Methods}

The methodology of our study follows the approach of \citet{willcoxetal2016}. 
We per
Summarize what we do a la:
In our simulations, we initialize a detonation once the 
deflagration front reaches a
characteristic DDT fuel density, which controls the degree of
expansion the star undergoes during the deflagration stage.

The comparison study of tradition C/O progentor WDs used the same
initial conditions for the realizations within the suite. Make the
point that the traditional models are simalted with the same code
and the same set if perturbations on the initial match head (that is
what we mean by a realization). The only difference is the different
nuclear buring to include the other Ne and the initial profile of
the 1-d model.

For much of the methodology to be described below, just say a few
words about what was done and cite \citet{willcoxetal2016}.

@@

This work was supported in part by the Department of Energy under
grant DE-FG02-87ER40317. The software used in this work was in part
developed by the DOE-supported ASC/Alliances Center for Astrophysical
Thermonuclear Flashes at the University of Chicago. Results in this
paper were obtained using the high-performance computing system at the
Institute for Advanced Computational Science at Stony Brook
University.  other grants...

>>>>

\section{Methodology}

The methodology of our study follows the approach of \citet{willcoxetal2016}. 
We performed a suite of two-dimensional simulations of supernova explosions
from hybrid progenitors and compared the results to a suite of simulations
of explosions from traditional C/O progenitor models performed with the
same simulation code and from similar initial conditions. The simulations assumed
the deflagration to detonation transition explosion paradigm, and the
transition densities were the same in both suites. We briefly our methodoly
here, and refer to the reader to previous work for additional details.
In particular, we use the same simulation instrument, a modified version
of the Flash Code, as \citet{willcoxetal2016} and we refer the reader there 
for a description of modifications for C-O-Ne burning and verification tests
of the burning module.

\subsection{Simulation Instrument}

The simulations of supernova explosions presented here were performed
with a customized version of the Flash code, originally developed
at the University of Chicago.
Flash is a parallel, adaptive mesh, multi-physics simulation code
developed first for nuclear astrophysics applications and subsequently
for high-energy-density applications~\cite{Fryxetal00,calder.curtis.ea:high-performance,
calder.fryxell.ea:on,flash_pragmatic,flash_evolution}. 
Flash has been applied to a variety of astropysical problems by a host
of researchers, and the version we apply differs from other versions
principally in the modules describing thermonuclear burning. 
specifically routines to describe
thermonuclear burning during both the deflagration and detonation
phases, as well as routines to describe the evolution of
the dynamic ash.

The need for a model flame in simulating thermonuclear supernovae
follows from the scales of the problem. At high densities, the
width of a laminar nuclear flame is $< 1\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$
while the radius of the white dwarf 
$\sim 10^9\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$.
Even with adaptive mesh refinement, whole-star simulations cannot simultaneousely
directly simulate the nuclear flame, thus requiring use of a model to
describe the burning on unresovable scales. For the deflagration phase,
our burning model propagates an artificially
broadened flame with an advection-diffusion-reaction (ADR)
scheme~\cite{Khok95,VladWeirRyzh06} via evolution of
a reaction progress variable to describe the consumption of C
and additional variables to describe the evolution of intermediate-mass
elements into the stastical quasi-equilibruim of the 
Si-group~\cite{ifk1981,khok1981,khok1983} and then into iron-group elements
(IGEs) in full nuclear statistical equilibrium (NSE).

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel and $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP
 described by~\cite{VladWeirRyzh06}, 
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\cite{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning~\cite{timmes92,Chamulak2007The-Laminar-Fla},
these are boosted to account for speed-up of burning due to
unresolved buoyancy and background
turbulence~\cite{Khok95,gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

The ADR scheme describes the consumption of C and the subsequent stages
are described by separate progress variables and
separate relaxation times derived from full nuclear network calculations
\cite{Caldetal07,townetal2016}. In both the quasi-equilibrium and full equilibrium, 
the creation of light elements by photodisintegration balances the creation of heavy 
elements by fusion, maintaining the equilibrium. The relative balance depends on
thermodynamic conditions, e.g.\ density and temperature, and hydrodynamic motion
during the explosion changes the thermodynamic conditions and thus the
balance. Electron capture also influences the evolution by shifting the binding 
energy of the material and thereby changing the temperature due to released energy, 
by changing the Fermi energy and thus the pressure, and by emitting neutrinos that 
escape and remove energy. Electron captures also neutronize the material, which 
produces more neutron rich iron-group material at the expense of \Ni. Like the 
input flame speed, the burning model includes tabulated rates for these effects 
from detailed NSE calculations ~\cite{SeitTownetal09}. Accordingly, the burning
model is able to describe dynamic evolution of the ash in addition to the 
stages of C-O-Ne burning.

The burning model also describes the detonation phase with progress variables.  
In this case, the model evolves thermally-activated burning with the actual
temperature-dependent rate for C consumption, which allow a propagating shock to
trigger burning, i.e.\ to propagate a detonation front. The propagating detonation 
is able to describe the same stages of C burning as the deflagration case, including 
the relaxation into NSE~\cite[and references therein]{townetal2016}. Finally,
we again note that the burning model was adapted for the case of burning in
hybrid white dwarfs. Direct numerical simulations of burning that guided 
the modifications, details of the modifications to the burning model, and
verification tests may be found in \citet{willcoxetal2016}.

@@

This work was supported in part by the U.S.\ Department of Energy under
grant DE-FG02-87ER40317 and in part by the U.S.\ National Science Foundation
via a supplement to grant AST-1211563. 
Support also came from the Data + Computing = Discovery summer program at
the Institute for Advanced Computational Science at Stony Brook University.
The software used in this work was developed in part by the DOE-supported 
ASC/Alliances Center for Astrophysical Thermonuclear Flashes at the 
University of Chicago. 
The source code used for
these studies is available as a package compatible with the
current Flash code from \url{http://astronomy.ua.edu/townsley/code.}
Results in this
paper were obtained using the high-performance computing system at the
Institute for Advanced Computational Science at Stony Brook
University.
